### 2016-09-27 text/ms.Rmd
Correct the denominator in percent change
All edits before first submission to Fish and Fisheries

```{r, echo=FALSE}
# Values to use in the paper:
load("values.rda") # generated by `../analysis/9-output-values.R`

# Includes: mean_sim, mean_ram, slope_sim, auc_sim, ram_stocks_n

@@

\noindent \textbf{Abstract:} Fishery managers must often reconcile conflicting
estimates of population status and trend. Superensemble models, commonly used in
climate and weather forecasting, may provide an effective solution. This
approach uses predictions from multiple models as covariates in an additional
"superensemble" model fitted to known data. We evaluated the potential for
ensemble averages and superensemble models ("ensemble methods") to improve
estimates of population status and trend for fisheries. We fit four widely
applicable data-limited models that estimate stock biomass relative to the
equilibrium biomass at maximum sustainable yield ($B/B_\mathrm{MSY}$). We
combined these estimates of recent fishery status and trends in $B/B_\mathrm{MSY}$
with four ensemble methods: an ensemble average and three superensembles (a
linear model, random forest, and boosted regression tree).  We trained our
superensembles on a simulated dataset of 5760 stocks and tested them with
cross-validation and against a global database of `r ram_stocks_n` stock
assessments. Ensemble methods substantially improved estimates of population
status and trend. Random forest and boosted regression trees performed the best
at estimating population status: accuracy improved `r
mean_sim$mach_vs_ind_mare_fold`%, rank-order correlation between predicted and
true status improved from `r mean_sim$ind_corr_range` to `r
mean_sim$mach_corr_range`, and bias (median proportional error) declined from
`r mean_sim$ind_mre_range` to `r mean_sim$mach_mre_range`.
We found similar improvements when predicting trend and when applying the
simulation-trained superensembles to catch data for global fish stocks.
Superensembles can optimally leverage multiple model predictions;
however, they must be tested, formed from a diverse set of accurate models, and
built on a dataset representative of the populations to which they are applied.

@@

Ensemble methods, and in particular the machine learning superensemble models
(random forest and GBM), generally improved estimates of status and trend over
any individual model (Fig. \ref{hexagon}e--h, Fig. \ref{scatter-sim-slope}e--h).
Compared to the individual models, machine learning superensembles improved
accuracy (median absolute proportional error) by
`r mean_sim_basic$mach_vs_ind_mare_fold`%, increased rank-order correlation from
`r mean_sim_basic$ind_corr_range` to `r mean_sim_basic$mach_corr_range`, and reduced bias
(median proportional error) from `r mean_sim_basic$ind_mre_range` to
`r mean_sim_basic$mach_mre_range` (Fig. \ref{performance}a).
These superensembles also generally had better ability to distinguish if
simulated stocks were above or below $B/B_\mathrm{MSY} = 0.5$ (Fig. \ref{roc-sim}).
Results were similar when predicting trend: compared to individual models,
machine learning superensembles improved accuracy by
`r slope_sim$mach_vs_ind_mare_fold`%, increased rank-order correlation from
`r slope_sim$ind_corr_range` to `r slope_sim$mach_corr_range`, and reduced
bias from `r slope_sim$ind_mre_range` to `r slope_sim$mach_mre_range` (Fig.
\ref{performance-sim-slope}).
The ensemble models that simply took a mean of the individual models ranked
slightly behind the best individual model for estimating fish stock status
(CMSY; Fig. \ref{performance}a) and had slightly lower correlation but higher
accuracy than the best individual model at predicting the trends of status
(SSCOM; Fig. \ref{performance-sim-slope}).

@@

Ensemble methods provide a useful approach to situations where environmental
resource management decisions must be made on the basis of multiple,
potentially contrasting estimates of status. Compared to individual models of
fish population status, ensemble methods were consistently the best or among
the best across three performance dimensions (accuracy, bias, and rank-order
correlation), two response variables (status and trend), two datasets
(simulated and global fisheries), and multiple ensemble methods (from a simple
average to machine learning superensembles). Our results suggest choosing a
superensemble model that allows for non-linear relationships, such as
machine learning methods; these models provided added insight into individual
model behavior and generally performed the best.

@@

# Acknowledgments

We thank members of Phase I of the working group "Developing new approaches to
global stock status assessment and fishery production potential of the seas"
who contributed to developing the data-limited methods and simulations used in
our analysis. We thank E. Jardim, F. Scott, and J.A. Hutchings for helpful
comments during the development of this project, and R.D. Methot for comments
on an earlier version of the manuscript. We thank the Gordon and Betty Moore
Foundation for funding the working group "Applying data-limited stock status
models and developing management guidance for unassessed fish stocks".

>>>>

```{r, echo=FALSE}
# Values to use in the paper:
load("values.rda") # generated by `../analysis/9-output-values.R`
if (mean_sim_basic$mach_mare_range == "0.32--0.32") mean_sim_basic$mach_mare_range <- 0.32
# Includes: mean_sim, mean_ram, slope_sim, auc_sim, ram_stocks_n

@@

\noindent \textbf{Abstract:} Fishery managers must often reconcile conflicting
estimates of population status and trend. Superensemble models, commonly used in
climate and weather forecasting, may provide an effective solution. This
approach uses predictions from multiple models as covariates in an additional
"superensemble" model fitted to known data. We evaluated the potential for
ensemble averages and superensemble models ("ensemble methods") to improve
estimates of population status and trend for fisheries. We fit four widely
applicable data-limited models that estimate stock biomass relative to the
equilibrium biomass at maximum sustainable yield ($B/B_\mathrm{MSY}$). We
combined these estimates of recent fishery status and trends in $B/B_\mathrm{MSY}$
with four ensemble methods: an ensemble average and three superensembles (a
linear model, random forest, and boosted regression tree).  We trained our
superensembles on 5760 simulated stocks and tested them with
cross-validation and against a global database of `r ram_stocks_n` stock
assessments. Ensemble methods substantially improved estimates of population
status and trend. Random forest and boosted regression trees performed the best
at estimating population status: inaccuracy (median absolute proportional error) decreased from
`r mean_sim_basic$ind_mare_range` to `r mean_sim_basic$mach_mare_range`,
rank-order correlation between predicted and
true status improved from `r mean_sim_basic$ind_corr_range` to
`r mean_sim_basic$mach_corr_range`, and bias (median proportional error) declined from
`r mean_sim_basic$ind_mre_range` to `r mean_sim_basic$mach_mre_range`.
We found similar improvements when predicting trend and when applying the
simulation-trained superensembles to catch data for global fish stocks.
Superensembles can optimally leverage multiple model predictions;
however, they must be tested, formed from a diverse set of accurate models, and
built on a dataset representative of the populations to which they are applied.

@@

Ensemble methods, and in particular the machine learning superensemble models
(random forest and GBM), generally improved estimates of status and trend over
any individual model (Fig. \ref{hexagon}e--h, Fig. \ref{scatter-sim-slope}e--h).
Compared to the individual models, machine learning superensembles decreased
inaccuracy (median absolute proportional error) from
`r mean_sim_basic$ind_mare_range` to `r mean_sim_basic$mach_mare_range`,
increased rank-order correlation from
`r mean_sim_basic$ind_corr_range` to `r mean_sim_basic$mach_corr_range`, and reduced bias
(median proportional error) from `r mean_sim_basic$ind_mre_range` to
`r mean_sim_basic$mach_mre_range` (Fig. \ref{performance}a).
These superensembles also generally had better ability to distinguish if
simulated stocks were above or below $B/B_\mathrm{MSY} = 0.5$ (Fig. \ref{roc-sim}).
Results were similar when predicting trend: compared to individual models,
machine learning superensembles decreased inaccuracy from
`r slope_sim$ind_mare_range` to `r slope_sim$mach_mare_range`,
increased rank-order correlation from
`r slope_sim$ind_corr_range` to `r slope_sim$mach_corr_range`, and reduced
bias from `r slope_sim$ind_mre_range` to `r slope_sim$mach_mre_range` (Fig.
\ref{performance-sim-slope}).
The ensemble models that simply took a mean of the individual models ranked
slightly behind the best individual model for estimating fish stock status
(CMSY; Fig. \ref{performance}a) and had slightly lower correlation but higher
accuracy than the best individual model at predicting the trends of status
(SSCOM; Fig. \ref{performance-sim-slope}).

@@

Ensemble methods provide a useful approach to situations where environmental
resource management decisions must be made on the basis of multiple,
potentially contrasting estimates of status. Compared to individual models of
fish population status, ensemble methods were consistently the best or among
the best across three performance dimensions (accuracy, bias, and rank-order
correlation), two response variables (status and trend), two datasets
(simulated and global fisheries), and multiple ensemble methods (from a simple
average to machine learning superensembles). Our results suggest choosing a
superensemble model that allows for non-linear relationships, such as
machine learning methods; these models provided added insight into individual
model behaviour and generally performed the best.

@@

# Acknowledgements

We thank members of Phase I of the working group "Developing new approaches to
global stock status assessment and fishery production potential of the seas"
who contributed to developing the data-limited methods and simulations used in
our analysis. We thank E. Jardim, F. Scott, and J.A. Hutchings for helpful
comments during the development of this project, and R.D. Methot for comments
on an earlier version of the manuscript. A.A. Rosenberg is president of MRAG
Americas Inc, a consulting company specializing in fisheries science. We thank
the Gordon and Betty Moore Foundation for funding the working group "Applying
data-limited stock status models and developing management guidance for
unassessed fish stocks".
